Electro-Drop Bouncing in Low-Gravity

### 2018-08-28 doc/paper/aiptemplate.tex
some paper updates

We investigate the dynamics of spontaneous jumps of water drops from electrically charged superhydrophobic dielectric substrates during a sudden step reduction in gravity level. In the brief free-fall environment of a drop tower, with a non-homogeneous external electric field arising due to dielectric surface charges (with surface potentials ~0.4-1.8 kV), body forces acting on the jumped drops are primarily supplied by polarization stress and Coulombic attraction instead of gravity. This electric body force leads to a drop bouncing behavior similar to well-known phenomena in 1-$g_0$, though occurring for much larger drops ($\sim$ 0.5 mL). We show a 1D model for the phenomenon, its scaling, and asymptotic estimates for drop time of flight in two regimes: at short-times close to the substrate when drop inertia balances Coulombic force due to net free charge and image charges in the dielectric substrate and at long-times far from the substrate when drop inertia balances free charge Coulombic force and drag. The drop trajectories are controlled primarily by the dimensionless electrostatic Euler number $\mathbb{E}\mbox{u}$, which is a ratio of inertia to electrostatic force. To experimentally determine values of $\mathbb{E}\mbox{u}$ we conduct a series of drop tower experiments where we observe the effects of drop volume, net free charge, and static surface potential of the superhydrophobic substrate on drop trajectories. We use a direct search optimization to obtain a Maximum Likelihood Estimate for drop net charge, as we do not measure it directly in experiment. For $\phi \mathbb{E}\mbox{u} /8 \pi > 1$ drops escape the electric field, where $\phi$ is a substrate to drop aspect ratio. However, we do not observe any escapes in our dataset.
\end{abstract}

@@

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands
\section{Spontaneous Drop Jump}
When a nonwetting, gravity-dominated sessile drop (e.g. a puddle), which is initially at rest on a surface in the Cassie-Baxter state suddenly undergoes a large step reduction in $\mathbb{B}\mbox{o} \equiv \frac{g R^2 \Delta \rho}{\gamma}$ it will spontaneously jump away from the surface, where $g$ is the acceleration, $R$ is the characteristic interfacial length scale, $\Delta \rho$ is the difference in densities across the interface (which simplifies to $\rho$, for large $\Delta \rho$, as in the case of an air-water interface), and where $\gamma$ is the surface tension. The spontaneous drop jump was first observed experimentally in the Soviet Union by Kirko \emph{et al.} \cite{kirko_phenomenon_1970} in 1970 for drops of mercury in hydrochloric acid, and later by Wollman \emph{et al.} in 2016 for water drops in air in a set of experiments conducted using drop towers \cite{wollman_more_2016}. The kinetic energy of the jump is supplied by the defect in free surface energy as the new minimum energy surface equilibrium has approximately constant curvature. The motive force is due to the inertia of internal flows which occur as the interface deflects under the suddenly lessened $\mathbb{B}\mbox{o}$. For drops with radial symmetry and sufficiently high initial $\mathbb{B}\mbox{o}$, inward radial capillary waves coalesce at the axis, leading to geysering and creation of satellite droplets by the Rayleigh-Plateau breakup of the geyser. In the case of smaller jumping drops the capillary waves are gradually damped by viscous forces.

@@

We have observed jumped drops to occasionally decelerate and return to the superhydrophobic surface, rebounding multiple times in the fashion of rigid bodies bouncing under 1-$g_0$. The forces at work in such situations are presumably electrostatic in origin. A time-lapsed composite image showing an example of the phenomenon is shown in Figure \ref{fig:bounce}, and schematically in Figure \ref{fig:bounce_time}.
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{bounce}
\caption{The trajectory of a 0.5 mL drop is captured in a composite image over a single bounce period ($\sim 1.25$ s) presented at $\approx 10$ Hz. The surface potential of the superhydrophobic dielectric is $\varphi_s = 1.25 \pm 0.41$ kV. \label{fig:bounce}}
\end{figure}

@@

We begin with some preliminary observations of the phenomenon:
\begin{itemize}
\item Observed maximum drop (de-)accelerations are on the order of $\sim$30 cm/s$^2$ for a range of drop volumes $0.03 \lessapprox V_d \lessapprox 0.5$ mL.
\item The water drops are attracted to regions of high electric field. The horizontal (surface plane parallel) components of the drop trajectory usually oscillate about some central position during the experiment (except in cases of nearly pure 1-D vertical translation). For especially small drops close to the spontaneous drop jump limit ($V_d \sim$ 0.01 mL) the drops do not jump, but translate across the surface in a rolling regime until either they reach a local maximum of the electric field, or until their motion is sufficiently damped by contact line hysteresis where pinning ultimately arrests their motion.
\item The drops appear to have net free charge. In cases of multiple simultaneous drop jumps the drops repel each other as they bounce or roll in orbital motion around regions of high field.
\item The magnitude of the drop trajectory maxima (apoapses) appear to be related to the drop volume (mass), and initial jump velocity (inertia), and to the electric field strength.
\end{itemize}

@@

The source of the net free charge on the drops is another issue. The drop charge could be due to the contact charging mechanism mentioned previously. For instance, in a 1996 paper, NASA flight engineer Don Pettit discusses the problem of low-gravity flow induced charging of liquids, resulting ultimately from contact charging phenomenon \cite{pettit_donald_flow_????}. However, a more likely mechanism for the drop charge is field-induced charging. Field-induced charging occurs due to breakup of a conductor having a field-induced dipole (e.g., a physical separation of charge). In our work this might occur when a drop is deposited on the charged surface by a grounded syringe. The metal syringe needle tip, and the liquid in the syringe itself are essentially a ground connection which is broken when the syringe is suddenly removed in the presence of an external electric field. Field-induced charging is at work in the famous Kelvin thunderstorm, and is applied in inkjet, and electrospay technologies, where in each case the breakup is by the Rayleigh-Plateau instability. Notably, in Pettit's aforementioned discussion of contact charging of liquids in low-g, he remarks on accidental electrostatic `hula-ing' of silicone oil drops when ejected from a syringe in the vicinity of a highly-charged polymer surface during an experiment conducted aboard STS-5 by mission specialist Joseph Allen \cite{pettit_donald_flow_????}. Depending on the (highly-variable) electrical conductivity of the silicone oil, and the material of the syringe used in the experiment, the charge could arise just as easily by field-induction as by contact charging. Relatedly, in a series of informal and somewhat whimsical experiments Pettit himself electrostatically orbited small water drops around a triboelectrically charged PTFE knitting needle while aboard ISS during expedition 30/31 \cite{stevenson_electrostatic_2015}. Again, in this case, the drop charge is likely field-induced.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{../figures/apparatus1.pdf}
\label{apparatus}
\caption{Schematic representation of drop jump with return and rebound from an electrically charged superhydrophobic substrate. The characteristic time and length scales $t_c$ and $y_c$ describe the time of flight and apoapse associated with the drop trajectory.}
\end{figure}

@@

In modeling the electrostatic force we begin with the standard electrohydrodynamic (EHD) approximation \cite{saville_electrohydrodynamics:_1997}. Under a DC electric field, we assume that the real part of the dielectric permittivity $\epsilon$, $\mbox{Re} \left( \epsilon \right) \approx  \mbox{constant}$. We also assume that electric currents are small enough that the effects of magnetic fields can be neglected. The validity of this assumption rests on the characteristic time scale $\tau_e = \epsilon /\sigma_e \ll 1$, where $\tau_e$ is the ratio of absolute dielectric permittivity $\epsilon = \kappa \epsilon_0$, to conductivity $\sigma_e$, of the medium, $\kappa$ is the relative dielectric permittivity, and $\epsilon_0$ is the vacuum permittivity. This characteristic time $\tau_e$ is also known as the relaxation time, and is a measure of how quickly the polarization of a dielectric responds to a change in electric field. Given the conductivity and permittivity in the limiting case of extremely-pure water ($ \epsilon \approx 80$, $\sigma_e = 18.2 \times 10^{6}$ $\Omega\mbox{cm}$) \cite{yatsuzuka_electrification_1994}, we estimate $\tau_e \approx 4 \times 10^{-6}$ s. The relaxation time for the common distilled water that is actually used in the drop experiments is undoubtedly shorter due to the presence of solvated ions. Neglecting the effects of an electric double layer on hydration of ions in the water or the ambient atmosphere due to the relatively massive size of the drops studied, the assumption of small relaxation time further implies that the free charge present in the drops will remain approximately constant during the typical time interval of a low-gravity experiment.

@@

By substituting Equations \ref{dep_force}, and \ref{image_force} into \ref{e_force} we have
\begin{eqnarray*}
 \mathbf{F}_E &=& q \mathbf{E} + \mathbf{F}_{DEP} + \mathbf{F}_I \\
 &=& q \mathbf{E} + \frac{k q^2}{16 \pi \epsilon_0 } y^{-2} \hat{\mathbf{j}} + 2 \pi R_d^3 \kappa_w \epsilon_0 K \nabla E^2, 
 \label{e_forces}
\end{eqnarray*}

By comparing DEP and Coulombic terms in Equation \ref{e_forces}, we note that a condition to neglect the DEP force is
\begin{eqnarray}
\frac{ \kappa_w \epsilon_0 K R_d^2 E_0}{q} \ll 1. \nonumber
\end{eqnarray}
As this condition likely prevails in our experiments we henceforth neglect the DEP force. There is some physical intuition to support this conclusion as well. The dielectric displacement $\mathbf{D} = \kappa \epsilon_0 \mathbf{E}$ of a water drop in air is high due to the large relative dielectric constant of water. This implies that the field strength within the drop is about 80 times smaller than in the surrounding medium. Thus it is not particularly inaccurate to model the dielectric volume of a drop as an equipotential conductive shell with zero field in its interior. As an aside, in treating the drop as an ideal conductor we note that in our specific case the electrostatic force is not a body force \emph{per se} as the electric field is acting on charges on the surface of the conductor.

Thus the 1-D governing equation becomes
\begin{equation}
 \label{gov_eqn_subs}
m y'' = - \frac{1}{2} C_D \rho A {y'}^2 - q E - \frac{k q^2}{16 \pi \epsilon_0} y^{-2},
\end{equation}
subject to
\begin{equation}
y(0) = R_d, \hspace{5 mm} \mbox{and} \hspace{5 mm} y'(0) = U_0 .
\end{equation}

@@

By taking Taylor series expansions in large and small $y$-limits we can intuit a bit about the behavior of this field. In the limit $y/L \ll 1$ Equation \ref{e_field} reduces to
\begin{equation}
\label{near_field}
E \approx \frac{\sigma}{4 \pi \epsilon_0} = E_0,
\end{equation}
where $E_0$ is the characteristic electric field. This field is constant and equivalent to the electric field due to an infinite plane of charge. In the limit of $y/L \gg 1$, Equation \ref{e_field} reduces to the familiar electric field due to a point charge
\begin{equation}
\label{far_field}
E \approx L^2 E_0 y^{-2}.
\end{equation}
Both regimes given by Equations \ref{near_field} and \ref{far_field} can be clearly seen in Figure \ref{fig:E0}.
\begin{figure}[h]
    \centering
    \def\svgwidth{\columnwidth}
    \input{../figures/E0.pdf_tex}
    \caption{A log-log plot of the magnitude of the dimensionless electric field $E$.\label{fig:E0}}
\end{figure}

\subsection{Scaling}
Equation \ref{gov_eqn_subs} is non-linear, non-homogeneous and must be solved numerically. Nevertheless we are curious if we can obtain approximate solutions to the equation analytically using perturbation methods. This is aided by non-dimensionalizing the governing equations, forming dimensionless groups, and determining whether any are particularly small or large, and whether any fortuitous opportunities for simplification appear. Introducing the scaled variables
\begin{equation}
 t^* = \frac{t}{t_c}, \hspace{10 mm} y^* = \frac{y}{y_c}, \end{equation}
where $y_c$ and $t_c$ are characteristic length and time scales respectively, and using the coordinate transformation $y(0) - R = 0$, the governing equation becomes
\begin{equation}
\label{pi_terms}
 {y^*}'' = - \mathbf{\Pi}_1 {y^*}'^2
- \mathbf{\Pi}_2 E^* ( {y^*} ) 
- \mathbf{\Pi}_3 \left( \mathbf{\Pi}_4 {y^*} + 1 \right)^{-2},
\end{equation}
subject to
\begin{equation*}
{y^*}(0) = 0, \hspace{5 mm} \mbox{and} \hspace{5 mm} {y^*}'(0) = \mathbf{\Pi}_5,
\end{equation*}
where we note the existence of several dimensionless groups
\begin{gather*}
\mathbf{\Pi}_1 = \frac{C_D \rho A y_c}{2 m}, \hspace{5 mm}
\mathbf{\Pi}_2 = \frac{q E_0 t_c^2}{m y_c}, \hspace{5 mm}
\mathbf{\Pi}_3 = \frac{k q^2 t_c^2}{16 \pi \epsilon_0 R_d^2 m y_c}, \hspace{5 mm}\\
\mathbf{\Pi}_4 = \frac{y_c}{R_d}, \hspace{5 mm}
\mathbf{\Pi}_5 = \frac{U_0 t_c}{y_c}.
\end{gather*}

\subsubsection{Inertial Electro-Image Limit}
In the limit of small $y$ and $t$ we expect inertia to scale with Coulombic and image forces. In this limit we can approximate the electric field as the constant $E_0$. One possible characteristic length scale is $y_c \sim R_d$, however this scale is overly restrictive with respect to time. With $y_c \sim U_0 t_c$ and picking $t_c$ such that the Coulombic force $\mathbf{\Pi}_3 \sim \mathcal{O}(1)$, the intrinsic scales are found such that
\[ t_c \sim \frac{m U_0}{q E_0}, \hspace{5 mm} \mbox{and} \hspace{5 mm} 
y_c \sim \frac{m U_0^2}{q E_0} .
\]
With these scales the governing equation \ref{pi_terms} becomes
\begin{equation}
{y^*}'' = -1 - \mathbb{I}\mbox{g} \left( \mathbb{E}\mbox{u}{y^*} + 1 \right)^{-2} , \label{img_limit}
\end{equation}
subject to
\begin{equation*}
{y^*}(0) = 0, \hspace{5 mm} \mbox{and} \hspace{5 mm} {y^*}'(0) = 1 .
\end{equation*}
with 
\[ \mathbb{I}\mbox{g} \equiv \frac{k q}{16 \pi \epsilon_0 R_d^2 E_0} = \mathbf{\Pi}_3, \hspace{5 mm}
\mathbb{E}\mbox{u} \equiv \frac{m U_0^2}{q E_0 R_d} = \mathbf{\Pi}_4 ,
\]
where the Image number $\mathbb{I}\mbox{g}$ is the ratio of image forces to the Coulombic force of the unperturbed field, and the electrostatic Euler number $\mathbb{E}\mbox{u}$ is the ratio of inertia to electrostatic force. With these dimensionless groups the intrinsic scales become
\[ t_c = \mathbb{E}\mbox{u} \frac{R_d}{U_0}, \hspace{5 mm} \mbox{and} \hspace{5 mm}
y_c = \mathbb{E}\mbox{u} R_d.
\]
We note that in a conservative system $\mathbb{E}\mbox{u}$ can be thought of as a ratio of energies: the kinetic energy $m U_0^2$ in the numerator and the electrostatic potential energy $q E_0 R_d$ in the denominator.

\subsubsection{Inertial Electro-Viscous Limit}
In the limit of large $y$ and $t$ we expect drop inertia to balance Coulombic force and drag. Here we approximate the electric field as $E \approx y_c^2 E_0 y^{-2}$. We choose the scaling $y_c \sim U_0 t_c$ and $\mathbf{\Pi}_3 \sim \mathcal{O}(1)$ for its combination of physical simplicity, few $\mathbf{\Pi}$ terms, and homogeneous initial conditions. The intrinsic scales for this case are given by
\[ t_c \sim \frac{R_d^2}{L^2} \frac{4 \pi m U_0}{q E_0} \hspace{5mm} \mbox{and} \hspace{5 mm} y_c \sim \frac{R_d^2}{L^2} \frac{4 \pi m U_0^2}{q E_0}.
\]
With this scaling the non-dimensional governing equation is 
\begin{eqnarray}
&{y^*}'' = - \mathbb{D}\mbox{g} \phi \mathbb{E}\mbox{u} {y^*}'^2 - \left( \phi \mathbb{E}\mbox{u} {y^*} + 1 \right)^{-2}, & \label{drag_limit}
\end{eqnarray}
subject to 
\begin{equation*}
{y^*}(0) = 0, \hspace{5 mm} \mbox{and} \hspace{5 mm} {y^*}'(0) = 1, 
\end{equation*}
where we call $\mathbb{D}\mbox{g}$ the drag number $\mathbb{D}\mbox{g} \equiv \frac{C_D \rho_a}{\rho_l} = \mathbf{\Pi}_1 \phi^{-1}{\mathbb{E}\mbox{u}}^{-1}$, and $\phi = 4 \pi \frac{R_d^2}{L^2}$ is a dimensionless ratio of length scales.

\subsection{Asymptotic Estimates}
The scalings of the equation of motion given by Equations \ref{img_limit} and \ref{drag_limit} are weakly non-linear differential equations in the sense that they reduce to linear equations as the parameter $\mathbb{E}\mbox{u} \rightarrow 0$. If we take $\mathbb{E}\mbox{u}$ to be a small parameter we can find an asymptotic approximation to the solution of the non-linear equation by means of a regular perturbation. In this case we use the naive expansion
\begin{equation}
{y^*}({t^*}) \sim y^*_0({t^*}) + \mathbb{E}\mbox{u} y^*_1({t^*}) + \mathbb{E}\mbox{u}^2 y^*_2({t^*}) \ldots \mathbb{E}\mbox{u}^n y^*_n ({t^*})  . \label{regular_pert}
\end{equation}

>>>>

We investigate the dynamics of spontaneous jumps of water drops from electrically charged superhydrophobic dielectric substrates during a sudden step reduction in gravity level. In the brief free-fall environment of a drop tower, with a non-homogeneous external electric field due to dielectric surface charges (with surface potentials ~0.4-1.8 kV), body forces acting on the jumped drops are primarily supplied by polarization stress and Coulombic attraction instead of gravity. This electric body force leads to a drop bouncing behavior similar to well-known terrestrial phenomena, though occurring for much larger drops ($\sim$ 0.5 mL). We provide a 1-dimensional model for the phenomenon, its scaling, and asymptotic estimates for drop time-of-flight in two regimes: at short-times close to the substrate when drop inertia balances Coulombic force due to net free charge and image charges in the dielectric substrate and at long-times far from the substrate when drop inertia balances free charge Coulombic force and drag. The drop trajectories are controlled primarily by the dimensionless electrostatic Euler number $\mathbb{E}\mbox{u}$, which is a ratio of inertia to electrostatic force. To experimentally determine values of $\mathbb{E}\mbox{u}$ we conduct a series of drop tower experiments where we observe the effects of drop volume, net free charge, and static surface potential of the superhydrophobic substrate on drop trajectories. We use a direct search optimization to obtain a Maximum Likelihood Estimate for drop net charge, as we do not measure it directly. For $\phi \mathbb{E}\mbox{u} /8 \pi > 1$ drops escape the electric field, where $\phi$ is a substrate-to-drop aspect ratio. However, we do not observe any escapes in our dataset since only drops where $\phi \mathbb{E}\mbox{u} /8 \pi < 1$ are achieved.
\end{abstract}

@@

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands
\section{Spontaneous Drop Jump}
When a nonwetting, gravity-dominated sessile drop (e.g. a puddle), initially at rest on a surface in the Cassie-Baxter state suddenly undergoes a large step reduction in Bond number $\mathbb{B}\mbox{o} \equiv \frac{g R^2 \Delta \rho}{\gamma}$ it will spontaneously jump away from the surface, where $g$ is the acceleration, $R$ is the characteristic interfacial length scale, $\Delta \rho$ is the difference in densities across the interface (which simplifies to $\rho$, for large $\Delta \rho$, as in the case of an air-water interface), and where $\gamma$ is the surface tension. The spontaneous drop jump was first observed experimentally in the Soviet Union by Kirko \emph{et al.} \cite{kirko_phenomenon_1970} in 1970 for drops of mercury in hydrochloric acid, and later by Wollman \emph{et al.} in 2016 for water drops in air in a set of experiments conducted using drop towers \cite{wollman_more_2016}. The kinetic energy of the jump is supplied by the defect in free surface energy as the new minimum energy surface equilibrium has approximately constant curvature. The motive force is due to the inertia of internal flows which occur as the interface deflects under the suddenly lessened $\mathbb{B}\mbox{o}$. For drops with radial symmetry and sufficiently high initial $\mathbb{B}\mbox{o}$, inward radial capillary waves coalesce at the axis, leading to geysering and creation of satellite droplets by the Rayleigh-Plateau breakup of the geyser. In the case of smaller jumping drops the capillary waves are quickly damped by viscous forces.

@@

We have observed jumped drops to occasionally decelerate and return to the superhydrophobic surface, rebounding multiple times in the fashion of rigid bodies bouncing under 1-$g_0$. The forces at work in such situations are presumably electrostatic in origin. A time-lapsed composite image showing an example of the phenomenon is shown in Figure \ref{fig:bounce} for a 0.5 mL drop, and schematically in Figure \ref{fig:bounce_time} for a 0.05 mL drop.
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{bounce}
\caption{The trajectory of a 0.5 mL drop is captured in a composite image over a single bounce period ($\sim 1.25$ s) presented at $\approx 10$ Hz. The surface potential of the superhydrophobic dielectric is $\varphi_s = 1.25 \pm 0.41$ kV. \label{fig:bounce}}
\end{figure}

@@

We begin with some preliminary observations of the phenomenon:
\begin{itemize}
\item Observed maximum drop (de-)accelerations are on the order of $\sim$30 cm/s$^2$ for a range of drop volumes $0.03 \lessapprox V_d \lessapprox 0.5$ mL.
\item The water drops are attracted to regions of high electric field. The horizontal (surface plane parallel) components of the drop trajectory usually oscillate about some central position during the experiment (except in cases of nearly pure 1-D vertical translation). For especially small drops close to the spontaneous drop jump limit (identified by Attari \emph{et al.} \citep{attari_puddle_2016} as $V_d \sim$ 0.01 mL) the drops do not jump, but translate across the surface in a rolling regime until either they reach a local maximum of the electric field, or until their motion is sufficiently damped by contact line hysteresis where pinning ultimately arrests their motion.
\item The drops appear to have net free charge. In cases of multiple simultaneous drop jumps the drops repel each other as they bounce or roll in orbital motion around regions of high field.
\item The magnitude of the drop trajectory maxima (apoapses) are related to the drop volume (mass), and initial jump velocity (inertia), and to the electric field strength.
\end{itemize}

@@

The source of the net free charge on the drops is another issue. The drop charge could be due to the contact charging mechanism mentioned previously. For instance, in a 1996 paper, NASA flight engineer Don Pettit discusses the problem of low-gravity flow induced charging of liquids, resulting ultimately from contact charging phenomenon \cite{pettit_donald_flow_????}. However, a more likely mechanism for the drop charge is field-induced charging. Field-induced charging occurs due to breakup of a conductor having a field-induced dipole (e.g., a physical separation of charge). In our work this might occur when a drop is deposited on the charged surface by a grounded syringe. The metal syringe needle tip and the liquid in the syringe itself are essentially a ground connection which is broken when the syringe is suddenly removed in the presence of an external electric field. Field-induced charging is at work in the famous Kelvin thunderstorm and is applied in inkjet and electrospay technologies, where in each case the breakup is by the Rayleigh-Plateau instability. Notably, in Pettit's aforementioned discussion of contact charging of liquids in low-g, he remarks on accidental electrostatic `hula-ing' of silicone oil drops when ejected from a syringe in the vicinity of a highly-charged polymer surface during an experiment conducted aboard STS-5 by Space Shuttle mission specialist Joseph Allen \cite{pettit_donald_flow_????}. Depending on the (highly-variable) electrical conductivity of the silicone oil, and the material of the syringe used in the experiment, the charge could arise just as easily by field-induction as by contact charging. Relatedly, in a series of informal and somewhat whimsical experiments Pettit himself electrostatically orbited small water drops around a triboelectrically charged PTFE knitting needle while aboard ISS during expedition 30/31 \cite{stevenson_electrostatic_2015}. Again, in this case, the drop charge is likely field-induced.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{../figures/apparatus3.pdf}
\label{apparatus}
\caption{Schematic representation of drop jump with return and rebound from an electrically charged superhydrophobic substrate. The characteristic time and length scales $t_c$ and $y_c$ describe the time of flight and apoapse associated with the drop trajectory. $L$ denotes the lengthscale characteristic of the charged superhydrophobic substrate.}
\end{figure}

@@

In modeling the electrostatic force we begin with the standard electrohydrodynamic (EHD) approximation \cite{saville_electrohydrodynamics:_1997}. Under a DC electric field, we assume that the real part of the dielectric permittivity $\epsilon$, $\mbox{Re} \left( \epsilon \right) \approx  \mbox{constant}$. We also assume that electric currents are small enough that the effects of magnetic fields can be neglected. The validity of this assumption rests on the characteristic time scale $\tau_e = \epsilon /\sigma_e \ll 1$, where $\tau_e$ is the ratio of absolute dielectric permittivity $\epsilon = \kappa \epsilon_0$, to conductivity $\sigma_e$, of the medium, $\kappa$ is the relative dielectric permittivity, and $\epsilon_0$ is the vacuum permittivity. This characteristic time $\tau_e$ is also known as the relaxation time, and is a measure of how quickly the polarization of a dielectric responds to a change in electric field. Given the conductivity and permittivity in the limiting case of extremely-pure water ($ \epsilon \approx 80$, $\sigma_e = 18.2 \times 10^{6}$ $\Omega\mbox{cm}$) \cite{yatsuzuka_electrification_1994}, we estimate $\tau_e \approx 4 \times 10^{-6}$ s. The relaxation time for the common distilled water that is actually used in the drop experiments is shorter due to the presence of solvated ions which tend to increase the conductivity. Neglecting the effects of an electric double layer on hydration of ions in the water or the ambient atmosphere due to the relatively massive size of the drops studied, the assumption of small relaxation time further implies that the free charge present in the drops will remain approximately constant during the typical time interval of a low-gravity experiment.

@@

By substituting Equations \ref{dep_force} and \ref{image_force} into \ref{e_force} we have
\begin{eqnarray}
 \mathbf{F}_E &=& q \mathbf{E} + \mathbf{F}_{DEP} + \mathbf{F}_I \nonumber \\
 &=& q \mathbf{E} + \frac{k q^2}{16 \pi \epsilon_0 } y^{-2} \hat{\mathbf{j}} + 2 \pi R_d^3 \kappa_w \epsilon_0 K \nabla E^2, 
 \label{e_forces}
\end{eqnarray}

By comparing DEP and Coulombic terms in Equation \ref{e_forces}, we note that a condition to neglect the DEP force is
\begin{eqnarray}
\frac{ \kappa_w \epsilon_0 K R_d^2 E_0}{q} \ll 1. \nonumber
\end{eqnarray}
We assume this conditions prevails in our experiments and henceforth neglect the DEP force. There is some physical intuition to support this conclusion as well. The dielectric displacement $\mathbf{D} = \kappa \epsilon_0 \mathbf{E}$ of a water drop in air is high due to the large relative dielectric constant of water. This implies that the field strength within the drop is about 80 times smaller than in the surrounding medium. Thus it is not particularly inaccurate to model the dielectric volume of a drop as an equipotential conductive shell with zero field in its interior. As an aside, in treating the drop as an ideal conductor we note that in our specific case the electrostatic force is not a body force \emph{per se} as the electric field is acting on charges on the surface of the conductor.

Thus the 1-D governing Equation \ref{e_forces}, becomes
\begin{equation}
 \label{gov_eqn_subs}
m y'' = - \frac{1}{2} C_D \rho A {y'}^2 - q E - \frac{k q^2}{16 \pi \epsilon_0} y^{-2},
\end{equation}
subject to
\begin{equation}
y(0) = R_d, \hspace{5 mm} \mbox{and} \hspace{5 mm} y'(0) = U_0 .
\end{equation}

@@

By taking Taylor series expansions in large and small $y$-limits we can intuit a bit about the behavior of this field. In the limit $y/L \ll 1$ Equation \ref{e_field} reduces to
\begin{equation}
\label{near_field}
E \approx \frac{\sigma}{4 \pi \epsilon_0} = E_0,
\end{equation}
where $E_0$ is the characteristic electric field. This field is constant and equivalent to the electric field due to an infinite plane of charge. In the limit of $y/L \gg 1$, Equation \ref{e_field} reduces to the familiar electric field due to a point charge
\begin{equation}
\label{far_field}
E \approx L^2 E_0 y^{-2}.
\end{equation}
Both regimes given by Equations \ref{near_field} and \ref{far_field} are identified in Figure \ref{fig:E0}.
\begin{figure}[h]
    \centering
    \def\svgwidth{\columnwidth}
    \input{../figures/E0.pdf_tex}
    \caption{A log-log plot of the magnitude of the dimensionless electric field $E$.\label{fig:E0}}
\end{figure}

\subsection{Scaling}
Equation \ref{gov_eqn_subs} is non-linear, non-homogeneous and must be solved numerically. Nevertheless we seek approximate solutions to the equation analytically using perturbation methods. Introducing the scaled variables
\begin{equation}
 t^* = \frac{t}{t_c}, \hspace{10 mm} y^* = \frac{y}{y_c}, \end{equation}
where $y_c$ and $t_c$ are characteristic length and time scales respectively, and using the coordinate transformation $y(0) - R = 0$, the governing equation becomes
\begin{equation}
\label{pi_terms}
 {y^*}'' = - \mathbf{\Pi}_1 {y^*}'^2
- \mathbf{\Pi}_2 E^* ( {y^*} ) 
- \mathbf{\Pi}_3 \left( \mathbf{\Pi}_4 {y^*} + 1 \right)^{-2},
\end{equation}
subject to
\begin{equation*}
{y^*}(0) = 0, \hspace{5 mm} \mbox{and} \hspace{5 mm} {y^*}'(0) = \mathbf{\Pi}_5,
\end{equation*}
where we note the existence of several dimensionless groups
\begin{gather*}
\mathbf{\Pi}_1 = \frac{C_D \rho A y_c}{2 m}, \hspace{5 mm}
\mathbf{\Pi}_2 = \frac{q E_0 t_c^2}{m y_c}, \hspace{5 mm}
\mathbf{\Pi}_3 = \frac{k q^2 t_c^2}{16 \pi \epsilon_0 R_d^2 m y_c}, \hspace{5 mm}\\
\mathbf{\Pi}_4 = \frac{y_c}{R_d}, \hspace{5 mm}
\mathbf{\Pi}_5 = \frac{U_0 t_c}{y_c}.
\end{gather*}

\subsubsection{Inertial Electro-Image Limit}
In the limit of small $y$ and $t$ we expect inertia to scale with Coulombic and image forces. In this limit we can approximate the electric field as the constant $E_0$. One possible characteristic length scale is $y_c \sim R_d$, however this scale is overly restrictive with respect to time. With $y_c \sim U_0 t_c$ and picking $t_c$ such that the Coulombic force $\mathbf{\Pi}_3 \sim \mathcal{O}(1)$, the intrinsic scales are found such that
\[ t_c \sim \frac{m U_0}{q E_0}, \hspace{5 mm} \mbox{and} \hspace{5 mm} 
y_c \sim \frac{m U_0^2}{q E_0} .
\]
With these scales the governing equation \ref{pi_terms} becomes
\begin{equation}
{y^*}'' = -1 - \mathbb{I}\mbox{g} \left( \mathbb{E}\mbox{u}{y^*} + 1 \right)^{-2} , \label{img_limit}
\end{equation}
subject to
\begin{equation*}
{y^*}(0) = 0, \hspace{5 mm} \mbox{and} \hspace{5 mm} {y^*}'(0) = 1 .
\end{equation*}
with 
\[ \mathbb{I}\mbox{g} \equiv \frac{k q}{16 \pi \epsilon_0 R_d^2 E_0} = \mathbf{\Pi}_3, \hspace{5 mm}
\mathbb{E}\mbox{u} \equiv \frac{m U_0^2}{q E_0 R_d} = \mathbf{\Pi}_4 ,
\]
where the Image number $\mathbb{I}\mbox{g}$ is the ratio of image forces to the Coulombic force of the unperturbed field, and the electrostatic Euler number $\mathbb{E}\mbox{u}$ is the ratio of inertia to electrostatic force. With these dimensionless groups the intrinsic scales become
\[ t_c = \mathbb{E}\mbox{u} \frac{R_d}{U_0}, \hspace{5 mm} \mbox{and} \hspace{5 mm}
y_c = \mathbb{E}\mbox{u} R_d.
\]
We note that in a conservative system $\mathbb{E}\mbox{u}$ can be thought of as a ratio of kinetic energy $m U_0^2$ and the electrostatic potential energy $q E_0 R_d$.

\subsubsection{Inertial Electro-Viscous Limit}
In the limit of large $y$ and $t$ we expect drop inertia to balance Coulombic force and drag. Here we approximate the electric field as $E \approx y_c^2 E_0 y^{-2}$. We choose the scaling $y_c \sim U_0 t_c$ and $\mathbf{\Pi}_3 \sim \mathcal{O}(1)$ for its combination of physical simplicity, few $\mathbf{\Pi}$ terms, and homogeneous initial conditions. The intrinsic scales for this case are given by
\[ t_c \sim \frac{R_d^2}{L^2} \frac{4 \pi m U_0}{q E_0} \hspace{5mm} \mbox{and} \hspace{5 mm} y_c \sim \frac{R_d^2}{L^2} \frac{4 \pi m U_0^2}{q E_0}.
\]
With this scaling the non-dimensional governing equation is 
\begin{eqnarray}
&{y^*}'' = - \mathbb{D}\mbox{g} \phi \mathbb{E}\mbox{u} {y^*}'^2 - \left( \phi \mathbb{E}\mbox{u} {y^*} + 1 \right)^{-2}, & \label{drag_limit}
\end{eqnarray}
subject to 
\begin{equation*}
{y^*}(0) = 0, \hspace{5 mm} \mbox{and} \hspace{5 mm} {y^*}'(0) = 1, 
\end{equation*}
where we call $\mathbb{D}\mbox{g}$ the drag number $\mathbb{D}\mbox{g} \equiv \frac{C_D \rho_a}{\rho_l} = \mathbf{\Pi}_1 \phi^{-1}{\mathbb{E}\mbox{u}}^{-1}$, and $\phi = 4 \pi \frac{R_d^2}{L^2}$ is a drop-to-substrate aspect ratio.

\subsection{Asymptotic Estimates}
The scalings of Equations \ref{img_limit} and \ref{drag_limit} are weakly non-linear differential equations in the sense that they reduce to linear equations as $\mathbb{E}\mbox{u} \rightarrow 0$. For $\mathbb{E}\mbox{u} \ll 1$ we find an asymptotic solution of the non-linear equation by means of a regular perturbation. In this case we use the naive expansion
\begin{equation}
{y^*}({t^*}) \sim y^*_0({t^*}) + \mathbb{E}\mbox{u} y^*_1({t^*}) + \mathbb{E}\mbox{u}^2 y^*_2({t^*}) \ldots \mathbb{E}\mbox{u}^n y^*_n ({t^*})  . \label{regular_pert}
\end{equation}
